---
title: Compile peptide optimisation lists
author: "Jonas D. Förster"
project:
    execute-dir: file
---

# Info

When new optimization data is created, it has to be checked for consistency and submitted into all the optimization files.
When new optimization data is meant to be read in, this is done in the form of a .csv file. For templates, use the files in optimized_oe/. Readin .csv files are put in optimized_oe/readin/ and will be integrated when running this script.
Also execute this whenever new precursors are added to the workflow (new peptides in !Peptide Lists/)

A backup of all the prior data will be put into optimized_oe/bkp.

In case of inconsistencies, errors will be generated. To recover, fix errors manually or consider deleting all data but !all_peptides.csv to ensure that there will be a single ground truth.

# Setup

```{r setup}
# code is outsourced to speed up interactive use of this file
source("!make_optimizations_oe_code.R")
eval(code$setup)
```

# Settings

```{r}
# any table in the "/readin" folder will be processed and integrated into the master list.
# By default, empty fields will not overwrite existing data in the master list.
# Set to TRUE to force overwriting all existing data according to what is in the readin table.
# Precursor data for precursors or full columns that don't exist in any readin tables are unaffected.
readin_force_overwrite <- FALSE
```

```{r}
# Set to true to overwrite all info with data from the master file !all_peptides.csv
only_use_all_peptide_table <- FALSE
```

```{r}
# the location of the optimizations folder
opt_folder <- "optimized_oe"
```

# Processing

```{r}
# Read in precursor master list to be able to ensure information on which precursors need to be in the database is available
# New precursors might have to be integrated
# Old precursors might have been deleted
eval(code$readin_precursors)

# precursors_df
```

```{r}
# Read in all optimization data
# All columns are verified for integrity upon readin
eval(code$readin_optimizations)

# opt_data_df
```

```{r}
# Extensive validation of all existing data is performed to ensure that everything is consistent between all tables
# A consolidated table then has each precursor only once
eval(code$consolidate_optimizations)

# opt_data_consolidated_df
```

```{r}
# List homeless precursors
# These are precursors for which optimization data is present, but which are for unknown reasons not in the precursor master list anymore.
eval(code$list_homeless_precursors)
```

```{r}
# Read all tables in the /readin folder and validate the data
# All new data is then integrated into the main optimization database
eval(code$integrate_readin_data)

# walk(readin_df_list, print)
# opt_data_consolidated_df
```

```{r}
eval(code$expand_for_write_out)

# opt_data_new_df
```

```{r}
eval(code$write_out_backup)
eval(code$write_out_data)
```

```{r include=FALSE}
cat0("Finished: ", date(), "\n")
```
