---
title: Read in peptide rt information
author: "Jonas D. Förster"
project:
  execute-dir: file
params:
  # The folder where the id_all_precursors.csv is located. Relative to the project root.
  detection_folder: prtc\55min PRM speclib
---

# Info

Submit iRT data for a set of peptides.
iRT data is anchored to the iRT master peptides defined in "!irt_master_list.txt".

Uses skyline id_all.csv (!Skyline Templates/id_all.skyr) export.
The skyline file should contain the iRT peaks and target peptide peaks.
The iRT peptide info is used to calculate the iRT value of the target peptides.
The target peptide iRT information is then submitted to the optimization database (.csv files).
Peptides may have multipe RTs submitted in one go. In that case both are stored and later scripts will usually implement windows covering all of them.

# Setup

```{r setup}
# code is outsourced to speed up interactive use of this file
source("!submit_rt_oe_code.R")
eval(code$setup)
```

# Settings

```{r}
# the location of the optimizations folder
opt_folder <- "optimized_oe"
```

# Processing

```{r}
# Read in the data of the skyline id_all.csv export
# Includes precursor information and especially the RT for all picked peaks
eval(code$readin_id_all)
```

```{r}
# Gather the RT for the iRT peptides, which will serve as the anchoring points.
# At least 2 of the iRT peptides need to be present in the submitted file or an error is thrown.
eval(code$gather_irt_rt)

# prtc_df
```

```{r}
# Gather the RT for all target peptides in the id_all.csv
# Peptides present in the file, for which no peak was set, will generate a warning.
# Such missing information will not overwrite any prior iRT data in the database.
eval(code$gather_target_rt)

# rt_df
```

```{r}
# Add the iRT specification to the targets
# The iRT info is specified as X:Y::Z, where X is the number of the iRT peptide eluting before, y is the number of the PRTC eluting after, and Z is the relative position between these two if 0<=Z<=1
# Z can also be <0 or >1, if the peptide elutes before or after all available iRT peptides.
eval(code$calc_irt_specification)

# rt_df
```

# Submit the data

```{r}
# Read in all optimization data, write the new data in, and save everything again, by writing all the relevant tables in `opt_folder`
# A backup of the old data is stored in `opt_folder`/bkp/
eval(code$read_opt_data)
eval(code$merge_new_opt_data)
eval(code$write_out_opt_data)

# opt_data_df
# opt_data_new_df
```

```{r include=FALSE}
cat0("Finished: ", date(), "\n")
```
